### 2016-03-02 doc/algorithm.tex

\section{Random Tree Search Algorithm for Finding NE}\label{sec:apx-alg}