### 2016-03-10 doc/introduction.tex
Modifying Introduction

\section{Introduction}
With myriads of data from social science and emergence of large scale social and economic networks, studying the evolutionary behavior of complex decision systems has become a major issue in recent years. In fact, game theory is one of the strong mathematical tools which has proven quite useful in capturing and modeling the complex nature of human decision making in social and economics networks. Despite numerous advances and contributions made in this field of study, there are a lot of barriers and limitations concerning computational issues of different games such as finding Nash equilibrium points or even approximation the equilibrium points of the system.

In this context, we consider in this paper a class of problems known as resource allocation problems. In general, resource allocation games are referred to a class of games in which a set of agents or players are competing for the same set of resources and has many applications in various areas such as load balancing, peer-to-peer systems, congestion games, web-caches, content management, and market sharing games among many others \cite{pollatos2008social,pacifici2012convergence,gopalakrishnan2012cache,goemans2006market,masucci2014strategic,ackermann2008impact,fabrikant2004complexity}. In fact, one of the important features of such games is to increase the reliability of the entire network with respect to customer needs and to improve the availability of the resources for users.

Typically, resource allocations games can be divided in two general groups: \textit{uncapacitated} and \textit{capacitated}. In uncapacitated allocation games the agents can keep even up to all the possible resources in their caches. However, there is an associated cost of keeping certain resources in the catch and this eliminates the posibility that every agent keep all the available resources in its catch. An instance of such games has been studied in \cite{chun2004selfish}, where the authors were able to fully characterize the set of equilibrium points. However, in the capaciteted case the situation could be much more complicated as the constraints on the capacities of the agents couples the actions of agents much more than in the uncapacitated case. As a result there is no comparable characterization of equilibrium points in capacitated selfish replication games.

In this paper we focus on an instance of capacitated resource allocation games with binary preferences as was introduced in \cite{gopalakrishnan2012cache}, where ``binary preferences" captures the behavioral pattern where players are equally interested in some objects. In the balance of this paper, our focus will be on such games, which for simplicity we refer to as CSR games. In fact, CSR games are defined in terms of a set of available resources for each player, where the players are allowed to communicate through an undirected communication graph. Such a communication graph identifies the access cost among the players, and the goal for each player is to satisfy his/her customers' needs with minimum cost.

The problem of finding an equilibrium for CSR games have been studied in \cite{gopalakrishnan2012cache,etesami2014pure}. It was shown in \cite{gopalakrishnan2012cache} that when the number of resources is 2, there exists a polynomial time algorithm $\mathcal{O}(n^3)$ to find an equilibrium, where $n$ is the number of players in the game. This result has been improved in \cite{etesami2014pure} to a linear time algorithm $\mathcal{O}(n)$ when the number of resources is bounded above by 5. Moreover, a quasi-polynomial approximation algorithm $\mathcal{O}(n^{\ln D})$ for approximating any Nash equilibrium of the system within a constant factor has been given in \cite{etesami2015approximation}, where $D$ denotes the diameter of the network. In this paper we consider CSR games over general undirected networks and devise a randomized algorithm based on random tree search to find the Nash equilibrium points of the system.

The paper is organized as follows. In Section~\ref{sec:game-model}, we introduce CSR games over general undirected networks and review some relevant existing results on this problem. In Section \ref{sec:main-I}, we  characterize the equilibrium points of the system as maximizers of a well-defined function and theoretically justify that randomization can be quite beneficial for maximizing such function under certain cases. Following this, in Section \ref{sec:apx-alg} we device a random tree search algorithm for maximize the objective function. Through numerous simulation results we show that this method can be indeed quite effective. We conclude the paper with identifying future directions of research in Section~\ref{sec:conclusion}.

\textbf{Notations}: 
For a positive integer $n$, we let $[n]:=\{1,2,\ldots,n\}$. We use $\mathcal{G}=([n], \mathcal{E})$ for an undirected underlying network with a node set $\{1,2,\ldots,n\}$ and an edge set $\mathcal{E}$. For any two nodes $i, j \in [n]$, we let $d_{\mathcal{G}}(i,j)$ be the graphical distance between them, that is, the length of a shortest path which connects $i$ and $j$. For an arbitrary node $i\in [n]$ and an integer $r\ge 0$, we define a ball of radius $r$ and center $i$ to be the set of all the nodes in the graph $\mathcal{G}$ whose graphical distance to the node $i$ is at most $r$, i.e., $B_{\mathcal{G}}(i,r)=\{x\in \mathcal{V}| d_{\mathcal{G}}(i,x)\leq r\}$. We denote the cardinality of a finite set $A$ by $|A|$.